Community Embeddings for Friend Suggestions

### 2020-10-12 conference_101719.tex
next: Visual Example
visual example done

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{mathtools}
\usepackage{commath}
\usepackage[style=numeric, backend=bibtex, language=english, maxbibnames=99]{biblatex}

@@

The hyper-parameter number of communities $K$ can reduce the quadratic summand $\abs{V}^2N$ by only considering users of the same community for friend suggestions. The reduction in time complexity through Algorithm \ref{alg:friend_sug_com} is only scalable when considering $K$ to be scale with the input graph's size. This assumption can be made when considering communities as groups of friends: with ten times more users, these users form ten times more communities.

>>>>

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{subcaption}
\usepackage[style=numeric, backend=bibtex, language=english, maxbibnames=99]{biblatex}

@@

The runtimes for Algorithm \ref{alg:friend_sug} and \ref{alg:friend_sug_com} when considering $K$ as constant or scaling w.r.t the graph $G = (V, E)$ are shown in Table \ref{tab:runtimes}. The number of friend suggestions $N$ is considered constant w.r.t the graph $G = (V, E)$.

\begin{table}
    \centering
    \caption{Algorithm \ref{alg:friend_sug} and \ref{alg:friend_sug_com} time complexity.}
    \label{tab:runtimes}
    \begin{tabular}{l l | c}
        \multirow{2}{*}{$K$ constant} & Algorithm \ref{alg:friend_sug}     & $O(\abs{E} + \abs{V}^2)$                          \\
                                      & Algorithm \ref{alg:friend_sug_com} & $O(\abs{E} + \abs{V}^2)$                          \\
        \hline
        \multirow{2}{*}{$K$ scaling}  & Algorithm \ref{alg:friend_sug}     & $O(K + \abs{V}K + \abs{E} + \abs{V}^2)$           \\
                                      & Algorithm \ref{alg:friend_sug_com} & $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2}{K})$ \\
    \end{tabular}
\end{table}

When the hyperparameter number of communities $K$ is considered non-constant with respect to the graph $G$, Algorithm \ref{alg:friend_sug_com} reduces the quadratic runtime in comparison to Algorithm \ref{alg:friend_sug_com}.
$K$ can therefore reduce the quadratic summand $\abs{V}^2N$ by only considering users of the same community for friend suggestions. The reduction in time complexity through Algorithm \ref{alg:friend_sug_com} is only scalable when considering $K$ to be scale with the input graph's size. This assumption can be made when considering communities as groups of friends: with ten times more users, these users form ten times more communities.

\section{Visual Example}

The small \citetitle{KarateClubDS} dataset will be used to visually underline the motivation for using node embedding and community memberships generated by ComE BGMM+VI for generating friend suggestions for person nodes on the karate club graph \cite{KarateClubDS}.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{images/karateclub/spring.png}
    \caption{\citetitle{KarateClubDS} plotted with networkx's spring layout \cite{KarateClubDS, networkx}.}
    \label{fig:karateclub_spring}
\end{figure}

Fig. \ref{fig:karateclub_spring} shows the 34 nodes and 78 edges of the \citetitle{KarateClubDS} dataset plotted as a graph with the Python network package networkx's spring layout.\cite{networkx} Spring layout simulates a force-directed representation of the network, by treating edges as springs pulling nodes together and treating nodes as repelling objects, sometimes called an anti-gravity force.\cite{networkx} The node classification, represented by node color, shown in Fig. \ref{fig:karateclub_spring} was obtained by running ComE with BGMM and VI with the hyperparameters presented in Table \ref{tab:params_karateclub}.

\begin{table}
    \centering
    \caption{Hyperparameters used for the Karateclub dataset.}
    \label{tab:params_karateclub}
    \begin{tabular}{ l | c | r }
        parameter                    & notation & value     \\
        \hline
        \hline
        number\_walks                & $\gamma$ & $10$      \\
        \hline
        walk\_length                 & $\ell$   & $80$      \\
        \hline
        representation\_size         & $D$      & $2$       \\
        \hline
        num\_workers                 & \empty   & $10$      \\
        \hline
        num\_iter                    & \empty   & $3$       \\
        \hline
        reg\_covar                   & \empty   & $0.00001$ \\
        \hline
        batch\_size                  & \empty   & $50$      \\
        \hline
        window\_size                 & $\zeta$  & $10$      \\
        \hline
        negative                     & $m$      & $5$       \\
        \hline
        lr                           & \empty   & $0.025$   \\
        \hline
        alpha                        & $\alpha$ & $0.1$     \\
        \hline
        beta                         & $\beta$  & $0.1$     \\
        \hline
        down\_sampling               & \empty   & $0.0$     \\
        \hline
        communities                  & $K$      & $5$       \\
        \hline
        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$ \\
    \end{tabular}
\end{table}

As can be seen in Figure \ref{fig:karateclub_spring}, the two communities in the \citetitle{KarateClubDS} dataset are correctly identified, despite the number of communities being initialized with $K=5$. Thanks to the Bayesian approach to community modeling and optimization through variational inference, three unused communities are dropped, which leaves the prominent two communities present in the dataset. The node classification obtained reflects the node classification published in the \citeyear{ComE} original ComE paper by \citeauthor{ComE}.\cite{ComE}

\subsection{Training and Testing Split}

\citetitle{KarateClubDS} is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded. The leave-one-out method is used in literature for evaluating recommender systems.\cite{Zhao2019ATS, He2017NeuralCF}

The three users with the highest degree are chosen as testing users: user 30, user 0, and user 32. For each testing user, a LOO-testing edge is chosen: (30, 23), (0, 1), and (32, 30). These three edges are spared for the testing dataset and omitted from the training graph,

\subsection{Embedding}

ComE BGMM+VI is run with the hyperparameters listed in Table \ref{tab:params_karateclub} on the training dataset. The resulting node and community embeddings are visualized as node embeddings colored by community membership and community embeddings plotted as ellipses in Fig. \ref{fig:karateclub_emb}.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{images/karateclub/embeddings_5.png}
    \caption{\citetitle{KarateClubDS} training dataset node and community embedding and node classification.}
    \label{fig:karateclub_emb}
\end{figure}

\subsection{Friend Suggesting}

With embeddings generated, top-$N$ friend suggestion recommendations are generated from the training node and community embeddings for the testing users. For each testing user, a ranked list of non-friend users from the same community is generated (Algorithm \ref{alg:friend_sug_com}). A test user's friend suggestion list is ordered in ascending order by the euclidean distance between the node embeddings. Table \ref{tab:karateclub_emb_dists} shows the test users' friend suggestion ranked lists.

\begin{table}
    \centering
    \caption{Testing users' friend suggestions ranked by node embedding distance.}
    \label{tab:karateclub_emb_dists}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User        & Distance   \\
            \hline
            \hline
            24          & $0.137181$ \\
            \hline
            25          & $0.214332$ \\
            \hline
            \textbf{23} & $0.543287$ \\
        \end{tabular}
        \caption{user 33}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User       & Distance   \\
            \hline
            \hline
            \textbf{1} & $0.322270$ \\
            \hline
            16         & $1.030336$ \\
        \end{tabular}
        \caption{user 0}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User        & Distance   \\
            \hline
            \hline
            25          & $0.039816$ \\
            \hline
            24          & $0.045994$ \\
            \hline
            27          & $0.182484$ \\
            \hline
            26          & $0.255288$ \\
            \hline
            28          & $0.356683$ \\
            \hline
            9           & $0.501317$ \\
            \hline
            \textbf{30} & $0.668306$ \\
        \end{tabular}
        \caption{user 32}
    \end{subfigure}
\end{table}

The testing users' removed edges make it to the lists successfully; therefore, the removed friends are in the same community as the test users and can be recommended through friend suggestions.

The top-$N$ friend suggestions, generated for the testing users with Algorithm \ref{alg:friend_sug_com}, are evaluated against the three removed test edges.
The ranked lists of friend-suggestions shown in Figure \ref{tab:karateclub_emb_dists} is used to determine top-$N$ friend suggestions for each testing user and evaluate the recommendations my the hit-rate metric. The LOO-testing edges are printed bold in Table \ref{tab:karateclub_emb_dists}. The resulting top-$N$ hit-rates are plotted against the number of suggestions $N$ in Fig. \ref{fig:karateclub_hitrates}.

\begin{figure}
    \centering
    \includegraphics[width=.4\textwidth]{images/karateclub/hit_rates.png}
    \caption{Hit-rates from top-$N$ recommendations on \citetitle{KarateClubDS}.}
    \label{fig:karateclub_hitrates}
\end{figure}

Fig. \ref{fig:karateclub_hitrates} shows promising results for using ComE to generate top-$N$ friend suggestions on the small \citetitle{KarateClubDS}. The achieved hit-rate starts at 33\% for $N=1$, reaches 67\% at $N=3$, and maxes out at 100\% for $N=7$; therefore, when presenting three friend suggestions to one test user, at least in $66\%$ of cases a friend suggestion would be of interest.
